Bandit Multiclass Linear Classification: Efficient Algorithms for the Separable Case

### 2019-01-21 ICML-submission/appendix-proofs-of-polynomial-approximation-theorems.tex
nn alg
add one paragraph to intro + change algorithm style

We prove the following lemma in this section.
\autoref{theorem:polynomial-approximation-2} immediately
follows from this lemma by considering $p' = p \cdot 2^{-s(s+1)rm+1})$
and algebra.

@@

It remains to prove part 4 of \autoref{theorem:polynomial-approximation-2}.
For any $i \in \{0,1,2,\dots,s\}$ and any $v \in \R^d$ such that $\norm{v} \le 1$
define multivariate polynomials
\begin{align*}
f_{i,v}(x) & = \frac{\ip{v}{x}}{\gamma} - 2^i \; , \\
q_v(x) & = P_s \left( \frac{\ip{v}{x}}{\gamma} \right) \; , \\
a_v(x) & = A_{s,r} \left( \frac{\ip{v}{x}}{\gamma} \right) \; , \\
b_v(x) & = B_{s,r} \left( \frac{\ip{v}{x}}{\gamma} \right) \; .
\end{align*}
Note that
\begin{multline*}
p(x) = \left[ \sum_{i=1}^m a_{v_i}(x) \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} b_{v_j}(x) \right]  \\
- \left(m - \frac{1}{2} \right) \prod_{j=1}^n b_{v_j}(x) \; .
\end{multline*}
We bound the norms of these polynomials. We have
$$
\norm{f_{i,v}}^2 = \norm{v}^2/\gamma^2 + 2^{2i} \le 2 \cdot 2^s \; .
$$
where we used that $1/\gamma \le 2^s$ and $\norm{v} \le 1$.
Since $q_v(x) = f_{i,v}(\frac{\ip{v}{x}}{\gamma}) \prod_{i=1}^s \left(f_{i,v}(\frac{\ip{v}{x}})\right)^2$,
using part 1 of \autoref{lemma:properties-of-norm-of-polynomials} we upper bound the norm of $q_v$
as
\begin{align*}
\norm{q_v}^2
& \le (2s+1)^{2s+1} \norm{f_{0,v}}^2 \prod_{i=1}^s \norm{f_{i,v}}^4 \\
& \le  (2s+1)^{2s+1} (2 \cdot 2^s)^{2s + 1} \; .
\end{align*}
Using parts 3 and 2 of \autoref{lemma:properties-of-norm-of-polynomials} we upper bound the norm of $a_v$ as
\begin{align*}
\norm{a_v}^2
& \le 2\norm{(q_v)^r}^2 + 2\norm{(q_{-v})^r}^2 \\
& \le 2 r^{r(2s+1)} \norm{q_v}^{2r} + 2 r^{r(2s+1)} \norm{q_{-v}}^{2r} \\
& \le 4 r^{r(2s+1)} \left((2s+1)^{2s+1} (2 \cdot 2^s)^{2s + 1} \right)^{2r} \\
& = 4 \left(2^{2s} r (4s+2)^2 \right)^{(2s+1)r} \; .
\end{align*}
The same upper bound holds for $\norm{b_v}^2$. Finally,
\begin{align*}
& \norm{p} \\
& \le \left[ \sum_{i=1}^m \norm{a_{v_i} \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} b_{v_j}} \right] + \left(m - \frac{1}{2} \right) \norm{\prod_{j=1}^m b_{v_j}} \\
& \le \left[ \sum_{i=1}^m m^{(s+1/2)rm} \norm{a_{v_i}} \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} \norm{b_{v_j}} \right] \\
& \qquad + \left(m - \frac{1}{2} \right) m^{(s+1/2)rm} \prod_{j=1}^m \norm{b_{v_j}} \\
& \le (2m-1/2) m^{(s+1/2)rm} \left(4 \left(2^{2s} r (4s+2)^2 \right)^{(2s+1)r} \right)^{m/2} \\
& = (2m-1/2) 2^m \cdot \left(2^{2s} rm (4s+2)^2 \right)^{(s+1/2)rm} \; .
\end{align*}

>>>>

We prove the following lemma in this section.
\autoref{theorem:polynomial-approximation-2} immediately
follows from this lemma by considering $p' = p \cdot 2^{-s(s+1)rm+1}$
and algebra.

@@

It remains to prove part 4 of \autoref{theorem:polynomial-approximation-2}.
For any $i \in \{0,1,2,\dots,s\}$ and any $v \in \R^d$ such that $\norm{v} \le 1$
define multivariate polynomials
\begin{align*}
f_{i,v}(x) & = \frac{\ip{v}{x}}{\gamma} - 2^i \; , \\
q_v(x) & = P_s \left( \frac{\ip{v}{x}}{\gamma} \right) \; , \\
a_v(x) & = A_{s,r} \left( \frac{\ip{v}{x}}{\gamma} \right) \; , \\
b_v(x) & = B_{s,r} \left( \frac{\ip{v}{x}}{\gamma} \right) \; .
\end{align*}
Note that
\begin{multline*}
p(x) = \left[ \sum_{i=1}^m a_{v_i}(x) \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} b_{v_j}(x) \right]  \\
- \left(m - \frac{1}{2} \right) \prod_{j=1}^n b_{v_j}(x) \; .
\end{multline*}
We bound the norms of these polynomials. We have
$$
\norm{f_{i,v}}^2 = \norm{v}^2/\gamma^2 + 2^{2i} \le 2 \cdot 2^s \; .
$$
where we used that $1/\gamma \le 2^s$ and $\norm{v} \le 1$.
Since $q_v(x) = f_{i,v}(\frac{\ip{v}{x}}{\gamma}) \prod_{i=1}^s \left(f_{i,v}(\frac{\ip{v}{x}}{\gamma})\right)^2$,
using part 1 of \autoref{lemma:properties-of-norm-of-polynomials} we upper bound the norm of $q_v$
as
\begin{align*}
\norm{q_v}^2
& \le (2s+1)^{2s+1} \norm{f_{0,v}}^2 \prod_{i=1}^s \norm{f_{i,v}}^4 \\
& \le  (2s+1)^{2s+1} (2 \cdot 2^s)^{2s + 1} \; .
\end{align*}
Using parts 3 and 2 of \autoref{lemma:properties-of-norm-of-polynomials} we upper bound the norm of $a_v$ as
\begin{align*}
\norm{a_v}^2
& \le 2\norm{(q_v)^r}^2 + 2\norm{(q_{-v})^r}^2 \\
& \le 2 r^{r(2s+1)} \norm{q_v}^{2r} + 2 r^{r(2s+1)} \norm{q_{-v}}^{2r} \\
& \le 4 r^{r(2s+1)} \left((2s+1)^{2s+1} (2 \cdot 2^s)^{2s + 1} \right)^{2r} \\
& = 4 \left(2^{2s} r (4s+2)^2 \right)^{(2s+1)r} \; .
\end{align*}
The same upper bound holds for $\norm{b_v}^2$. Finally,
\begin{align*}
& \norm{p} \\
& \le \left[ \sum_{i=1}^m \norm{a_{v_i} \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} b_{v_j}} \right] + \left(m - \frac{1}{2} \right) \norm{\prod_{j=1}^m b_{v_j}} \\
& \le \left[ \sum_{i=1}^m m^{(s+1/2)rm} \norm{a_{v_i}} \prod_{\substack{j ~:~ 1 \le j \le m \\ j \neq i}} \norm{b_{v_j}} \right] \\
& \qquad + \left(m - \frac{1}{2} \right) m^{(s+1/2)rm} \prod_{j=1}^m \norm{b_{v_j}} \\
& \le (2m-1/2) m^{(s+1/2)rm} \left(4 \left(2^{2s} r (4s+2)^2 \right)^{(2s+1)r} \right)^{m/2} \\
& = (2m-1/2) 2^m \cdot \left(2^{2s} rm (4s+2)^2 \right)^{(s+1/2)rm} \; .
\end{align*}

### 2019-01-21 ICML-submission/introduction.tex
add one paragraph to intro + change algorithm style

\begin{protocol}[h]
\caption{\textsc{Online Multiclass Linear Classification with Bandit Feedback}
\label{algorithm:game-protocol}}
\begin{algorithmic}[1]
{
\REQUIRE{Number of classes $K$, number of rounds $T$.}
\FOR{$t=1,2,\dots,T$}
\STATE Adversary chooses example $(x_t, y_t)$, where $x_t$ is revealed to the learner.
\STATE{Predict class label $\widehat y_t \in \{1,2,\dots,K\}$.}
\STATE{Observe feedback
$z_t = \indicator{\widehat y_t \neq y_t} \in \{0,1\}$
%$z_t \in \{0,1\}$, \\ \qquad where $z_t = \begin{cases}
%0 & \text{if $\widehat y_t = y_t$,} \\
%1 & \text{if $\widehat y_t \neq y_t$.}
%\end{cases}$
}
\ENDFOR
}
\end{algorithmic}
\end{protocol}

@@

For the case when the examples are weakly linearly separable, we construct a
non-linear feature mapping $\phi$ and associated positive definite kernel
$k(x,x')$ that makes the examples \emph{strongly} linearly separable. We then
use the kernelized version of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} for
the strongly separable case (Algorithm~\ref{algorithm:kernelized}). The kernel
$k(x,x')$ corresponding to the feature mapping $\phi$ has a simple explicit
formula and can be computed in $O(d)$ time, making
Algorithm~\ref{algorithm:kernelized} computationally efficient. For details on
kernel methods see e.g.~\citet{Scholkopf-Smola-2002} or
\citet{Shawe-Taylor-Cristianini-2004}.

>>>>

\begin{protocol}[h]
\caption{\textsc{Online Multiclass Linear Classification with Bandit Feedback}
\label{algorithm:game-protocol}}
\textbf{Require:} Number of classes $K$, number of rounds $T$.\\
\For{$t=1,2,\dots,T$}{
     Adversary chooses example $(x_t, y_t)$, where $x_t$ is revealed to the learner.\\
     Predict class label $\widehat y_t \in \{1,2,\dots,K\}$.
     Observe feedback
          $z_t = \indicator{\widehat y_t \neq y_t} \in \{0,1\}$
%$z_t \in \{0,1\}$, \\ \qquad where $z_t = \begin{cases}
%0 & \text{if $\widehat y_t = y_t$,} \\
%1 & \text{if $\widehat y_t \neq y_t$.}
%\end{cases}$
}

\end{protocol}

@@

For the case when the examples are weakly linearly separable, it was open for a long 
time whether there exist \textit{efficient} algorithms with finite mistake 
bound~\cite{Kakade-Shalev-Shwartz-Tewari-2008, Beygelzimer-Orabona-Zhang-2017}. 
Furthermore, in~\cite{Kakade-Shalev-Shwartz-Tewari-2008}, they ask the following question:
is there \textit{any} algorithm with a finite mistake bound that has no explicit dependence 
on the dimensionality of the feature vectors? We answer both questions affirmatively by
providing an efficient algorithm with finite dimensionless mistake bound 
(Algorithm~\ref{algorithm:kernelized}) (Note: the second question was previously 
answered by~\cite{Daniely-Helbertal-2013} with an inefficient algorithm).

The strategy used in Algorithm~\ref{algorithm:kernelized} is to construct a
non-linear feature mapping $\phi$ and associated positive definite kernel
$k(x,x')$ that makes the examples \emph{strongly} linearly separable in a higher-dimensional space. 
We then use the kernelized version of
Algorithm~\ref{algorithm:algorithm-for-strongly-linearly-separable-examples} for
the strongly separable case (Algorithm~\ref{algorithm:kernelized}). The kernel
$k(x,x')$ corresponding to the feature mapping $\phi$ has a simple explicit
formula and can be computed in $O(d)$ time, making
Algorithm~\ref{algorithm:kernelized} computationally efficient. For details on
kernel methods see e.g.~\citet{Scholkopf-Smola-2002} or
\citet{Shawe-Taylor-Cristianini-2004}.
